### 2014-02-07 Beacon P3 Proposal/main.tex
Added GE to BEACON proposal.

Real world applications of evolutionary optimization are often
hindered by the need to determine problem specific parameter
settings.  
Parameters can have a significant impact on search efficiency~\cite{goldberg:1991:gasize}
and finding good settings can be very time consuming~\cite{grefenstette:1986:optimalga}.
While some previous methods have reduced~\cite{Back:1992:selfadapt}
or removed~\cite{pelikan:2004:parameterlesshboa} the
need for parameter tuning, many do so by trading efficiency for
general applicability.  The Parameter-less Population Pyramid (P3)
is an evolutionary technique that requires \textbf{no parameters}
and is still broadly effective. P3 strikes a balance between
continuous integration of diversity and exploitative elitist
operators, allowing it to solve easy problems quickly and hard
problems eventually.  Work currently under review for publication
compared P3 with three optimally tuned state-of-the-art algorithms~\cite{thierens:2013:ltgahiff,doerr:2013:lambdalambda}.
Across seven problem classes P3 always found the
optimum at least a constant factor \emph{\textbf{faster}} than all comparisons.
More importantly, on the three most challenging problem classes,
P3 had a lower order of computational complexity as measured by
evaluations.  Based on over one trillion evaluations, these results
suggest P3 has wide applicability to a broad class of
problems.  The next step for P3 is to determine if those results carry over into real world
applications.  \textbf{Therefore, the primary focus of this project is to integrate
P3 into X-TOOLSS in order to (1) allow non-experts easy access to P3 and
(2) gather evidence of P3's ability to optimize real world problems.}

@@

Beyond accessibility, X-TOOLSS provides an invaluable potential to P3 for academic data collection.
The framework's optimization analysis tools have been refined through continuous multi-user use, and
will likely lead to further insights into P3's optimization.  Furthermore, as X-TOOLSS has a number
of built in evolutionary optimizers, it provides a perfect device for gathering comparative results
as P3 is applied to new and challenging problems.  Once integrated, non-computer scientists can utilize
X-TOOLSS to compare P3 directly with existing evolutionary techniques.  This can provide the foundation
for collaborative research between industry and academia, with industry gaining high quality solutions
to challenging problems and academia gathering data on optimization techniques.

@@

@@

Once collected, P3's ability to tackle real problems will provide either further evidence for its usefulness
to people outside of the evolutionary optimization community or as a means for further improving P3.
In either case it will help drive the direction of future research in parameter-less evolutionary optimization.

@@

Beyond the direct benefits gained by NASA and X-TOOLSS users, better understanding of P3
can improve evolutionary optimization in general.  As mentioned previously, evolutionary
methods are already able to produce human competitive designs, with enough successes each
year to hold an annual competition.\footnote{\url{http://www.genetic-programming.org/combined.php}}
These results impact everything from space exploration to cancer screenings, software error correction
to protein structure prediction.

>>>>

Real world applications of evolutionary optimization are often
hindered by the need to determine problem specific parameter
settings.  
Parameters can have a significant impact on search efficiency~\cite{goldberg:1991:gasize}
and finding good settings can be very time consuming~\cite{grefenstette:1986:optimalga}.
While some previous methods have reduced~\cite{Back:1992:selfadapt}
or removed~\cite{pelikan:2004:parameterlesshboa} the
need for parameter tuning, many do so by trading efficiency for
general applicability.  The Parameter-less Population Pyramid (P3)
is an evolutionary technique that requires \textbf{no parameters}
and is still broadly effective. P3 strikes a balance between
continuous integration of diversity and exploitative elitist
operators, allowing it to solve easy problems quickly and hard
problems eventually.  Work currently under review for publication
compared P3 with three optimally tuned state-of-the-art algorithms~\cite{thierens:2013:ltgahiff,doerr:2013:lambdalambda}.
Across seven problem classes P3 always found the
optimum at least a constant factor \emph{\textbf{faster}} than all comparisons.
More importantly, on the three most challenging problem classes,
P3 had a lower order of computational complexity as measured by
evaluations.  Based on over one trillion evaluations, these results
suggest P3 has wide applicability to a broad class of
problems.  \textbf{Therefore, the primary focus of this project is to integrate
P3 into X-TOOLSS in order to (1) allow non-experts easy access to P3,
(2) improve P3's ability to optimize real world problems and (3) expand
P3 to new evolutionary domains utilizing Genetic Programming.}

@@

Beyond accessibility, X-TOOLSS provides an invaluable potential to P3 for academic data collection.
The framework's optimization analysis tools have been refined through continuous multi-user use, and
will likely lead to further insights into P3's optimization.  Furthermore, as X-TOOLSS has a number
of built in evolutionary optimizers, it provides a perfect device for gathering comparative results
as P3 is applied to new and challenging problems.  Once integrated, non-computer scientists can utilize
X-TOOLSS to compare P3 directly with existing evolutionary techniques.  This can provide the foundation
for collaborative research between industry and academia, with industry gaining high quality solutions
to challenging problems and academia gathering data on how to improve optimization techniques.

@@

P3 is currently limited to evolving discrete value encoded linear genomes. 
Therefore we propose to leverage Grammatical Evolution~\cite{oneill:2001:grammatical}
to give P3 the ability to evolve entire programs.  Grammatical Evolution (GE) is a subset
of Genetic Programming which utilizes a non-trivial genotype to phenotype mapping, converting
binary strings into full fledged programs.  To apply GE, a domain expert first creates a
grammar which defines the structure of all possible useful programs.  When evaluating a
solution, bits are read from the individual to determine which grammar expansion rules to perform,
deterministically creating a program.  This is an ideal transition for P3 as from an
evolutionary standpoint GE applications will be no different than binary string applications as
only the evaluation step will have to change.

@@

Once collected, P3's ability to tackle real problems will provide either further evidence for its usefulness
to people outside of the evolutionary optimization community or as a means for further improving P3.
In either case it will help drive the direction of future research in parameter-less evolutionary optimization.
Expanding P3 to evolve entire programs will also provide interesting insight into the challenging domain
of Genetic Programming.

@@

Beyond the direct benefits gained by NASA and X-TOOLSS users, better understanding of P3
can improve evolutionary optimization in general.  As mentioned previously, evolutionary
methods are already able to produce human competitive designs, with enough successes each
year to hold an annual competition.\footnote{\url{http://www.genetic-programming.org/combined.php}}
These results impact everything from space exploration to cancer screenings, software error correction
to protein structure prediction.  Evolution of programs can help P3 create truly innovative solutions to complex problems.
